---
title: "Week 4 Activity"
author: 'Group 5: Yan Chen, Zhilin He, Hao He, Tianshi Wang'
date: "9/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(datasets)
```

### Part 1 (dplyr)

In the part 1 of the activity you will write or interpret short dplyr pipes
that explore or analyze the `Orange` data from R's `datasets` package. The data
which contains 35 rows and 3 columns recording the growth
of orange trees.  The three columns in the data set are:

  - `Tree`: an ordered factor, identifying individual trees, 
  - `age`: a numeric vector giving the number of days 
    since the tree was planted,
  - `circumference`: a numeric vector recording the circumference of the 
    trunk in mm. 

You can read more at `?Orange`. 

1. Write a dplyr pipe to determine the number of
observations per tree.

```{r q1}
data("orange")
view(Orange)

num_obs = Orange %>% group_by(Tree) #%>% summarise(n(Tree))

num_obs
```

2. Write a dplyr pipe to change the units of age to "years" 
and circumference to "cm".

```{r q2}

Orange1 = Orange %>% transform(age = age/365, circumference = circumference/10)
Orange %>% transform(age = age/365, circumference = circumference/10)

```

 
3. Write a dplyr pipe to add a column assigning a
z-score to each tree, centered around the mean for all trees at a given age.

```{r q3}
Orange %>% group_by(age) %>% 
  summarise(mean = mean(circumference), sd = sd(circumference)) %>% 
  left_join(Orange, by = "age") %>%
  mutate(zscore = (circumference-mean)/sd) %>% select(-c(mean,sd))


```

 
4. Write a dplyr pipe to calculate the average rate of growth
(cm/year) across all trees between between age 0 (assume circumference = 0) and
the final measured age for each tree.

```{r q4}
Orange1 %>% mutate(avg_age = circumference/age)

Orange_r = Orange1 %>% mutate(avg_rate = circumference/age)
mean(Orange1$avg_rate)

```

5. Describe the result of the following pipe in 1-3 sentences. Your
    *description* should touch on both the rows and columns and also describe
    a substantive question addressed by the result. 

```{r}
 Orange %>%
   group_by(Tree) %>%
   mutate( 
     new_growth = c(circumference[1], diff(circumference)),
     elapsed = c(age[1], diff(age))
   ) %>%
   group_by(age) %>% 
   summarize( 
     avg_rate = mean(new_growth / elapsed),
     se = sd( new_growth / elapsed ) / sqrt(n()) 
   )
```

**Description:**
The result is a 7 by 3 data.frame which contains three rows: age, avg_rate, se.
For each row, it represents the average growth of tree and the standard deviation of rate for the given age group.
